Discovering cell type motifs in connectomics data

### 2013-12-05 docs/paper.tex
bib changes
some framing changes, still need to, you know, really work on text added hyperparams

@@

\todo{Citations}
\todo{Log score plots for real data} 
\todo{Fix vector notation}
\todo{Get feedback on joint model score}
\todo{Find other irm paper citation}
% \begin{affiliations}
%   \item Something Something
%   \item Something Something
% \end{affiliations}

\begin{abstract}
  New techniques produce massive data about neural connectivity,
  necessitating new analysis methods to discover underlying
  structure. There is strong evidence for microcircuitry, where the
  probability of connections depends on the type of pre- and
  post-synaptic neuron and their distance. Here we thus developed a
  nonparametric Bayesian technique that identifies type and circuitry
  motifs from the connection data. We show that the approach recovers
  known neuron types in the retina, reveals interesting structure in
  the nervous system of c. elegans, and automatically discovers the
  structure of microprocessors. Our approach is a first step towards
  condensing connectomics data into meaningful, human readable
  structure.
\end{abstract}

@@

The brain consists of multiple kinds of neurons, each of which is
hypothesized to have a specific role in the overall
computation. Neuron types differ in many ways, e.g. chemical or
morphological, but they also differ in the way they connect to one
another. In fact, the idea of well defined local connectivity motifs
has been prominent in many areas, ranging from sensory (e.g. retina,
\autocite{} via processing (e.g. neocortex) to movement (e.g. spinal cord)
\autocite{}. Indeed, such motifs also exists in human-made computing
circuits. It remains an important challenge to develop algorithms to
use anatomical data, e.g. connectomics, to back out underlying
connectivity motifs.

The discovery of structure is a crucial aspect of network
science. Early approaches focused on global graph properties, such as
the types of scaling present in the network \autocite {WattsStrogatz1998}
.  While this approach leads to an understanding of global
network, more recent work aims at identifying repeat patterns, or
“motifs” in networks. \todo{There's the Science2002 paper by Milo et. al. that explicitly looks at 'motifs'} A motif is a set of nodes with conserved
relations that appears multiple times in a network, for example a set
of three nodes that are each connected to one another. When it comes
to neuroscience, we might expect that a motif consists of the kinds of
cells that jointly make up a microcircuit along with their distance
and cell-type dependent connectivity. The same motif should then
appear in multiple locations in the nervous system. Combining
motif-discovery ideas with specific knowledge about the kinds of
motifs in the nervous system promises better structure discovery.

@@

To show the applicability of our method to other connectome-style
datasets, we obtained the spatial location and interconnectivity of a
classic microprocessor, the MOS Technology 6502 used in the original
Apple II, Atari, and Commodore 64 \autocite{datasource}. We identified a
region of the processor with complex but known structure containing
the primary general-purpose registers X, Y, and S (figure).

Our algorithm identifies areas of spatial homogeneity 
[mirror structure]

@@

For probabilistic models like ours, no known solution exists to
exactly find the most probable parsing of the neurons into cell-types
and connectivity patterns [homogenize these two wordings across the
paper]. We employ a collection of Markov-chain Monte carlo techniques
[reference methods] but while different initializations converge to
similar ultimate values, we can never realistically obtain the global
optimum. There are a broad range of techniques that may offer better
approximations to the global optimum \autocite{Wanga2012,FritzJonathan} and future work could
adapt them to the problem we addressed here.

@@

Our model is a extension of infinite stochastic block models
\autocite{Kemp2006a, other guys} to incorporate spatial constraints.

\todo{infinite subset relaitonal model has good reults}

@@

\subsection{Extension to multiple graphs}
\todo{derivation of extension to multiple graphs}

\subsection{Hyperprior grids}
\todo{Plot hyperparameter Graphs}

>>>>

\renewcommand{\vec}[1]{\mathbf{#1}}

@@

\todo{Citations}
\todo{Log score plots for real data} 
\todo{Hierarhcical structure?}
% \begin{affiliations}
%   \item Something Something
%   \item Something Something
% \end{affiliations}

\begin{abstract}
  New techniques produce massive data about neural connectivity,
  necessitating new analysis methods to discover underlying
  structure. There is strong evidence for microcircuitry, where the
  probability of connections depends on the types of pre- and
  post-synaptic neurons and their distance. Here we developed a
  nonparametric Bayesian technique that identifies neuron type and circuitry
  patterns from the connection data. We show that the approach recovers
  known neuron types in the retina, reveals interesting structure in
  the nervous system of c. elegans, and automatically discovers the
  structure of microprocessors. Our approach is a first step towards
  condensing connectomics data into meaningful, human readable
  structure.
\end{abstract}

@@

The brain consists of multiple kinds of neurons, each of which is
hypothesized to have a specific role in the overall
computation. Neuron types differ in many ways, e.g. chemical or
morphological, but they also differ in the way they connect to one
another. In fact, the idea of well defined local connectivity patterns
has been prominent in many areas, ranging from sensory (e.g. retina,
\autocite{} via processing (e.g. neocortex) to movement (e.g. spinal
cord) \autocite{}. Indeed, such repeated patterns also exists in
human-made computing circuits. It remains an important challenge to
develop algorithms to use anatomical data, e.g. connectomics, to back
out underlying connectivity motifs.

The discovery of structure is a crucial aspect of network
science. Early approaches focused on global graph properties, such as
the types of scaling present in the network \autocite
{WattsStrogatz1998} .  While this approach leads to an understanding
of global network, more recent work aims at identifying small-scale
repeat patterns, or “motifs” in networks. \todo{There's the
  Science2002 paper by Milo et. al. that explicitly looks at 'motifs'}
A motif is a set of nodes with conserved relations that appears
multiple times in a network, for example a set of three nodes that are
each connected to one another.

When it comes to neuroscience, we might
expect that a motif consists of the kinds of cells that jointly make
up a microcircuit along with their distance and cell-type dependent
connectivity. The same motif should then appear in multiple locations
in the nervous system. Combining motif-discovery ideas with specific
knowledge about the kinds of motifs in the nervous system promises
better structure discovery.

@@

To show the applicability of our method to other connectome-style
datasets, we obtained the spatial location and interconnectivity of a
transistors in a classic microprocessor, the MOS Technology 6502 (used
in the Apple II) \autocite{datasource}. We identified a region of the
processor with complex but known structure containing the primary
general-purpose registers X, Y, and S (figure). Computer architects
use common patterns of transistors when designing circuits, 
with each transistor having a ``type'' in the circuit.

Our algorithm identifies areas of spatial homogeneity that mirror the
kown structure in the underlying architectural circuit, segementing
transistor groups. One group contains the ``clocked'' transistors,
one group with their source connected to ground, and one with their
drain connected to ground. The repeat patterns of spatial connectivity
are visible in figure~\ref{fig:microprocessor}a,

@@

For probabilistic models like ours, no known solution exists to
exactly find the most probable parsing of the neurons into cell-types
and connectivity patterns . We employ a collection of Markov-chain
Monte carlo techniques [reference methods] but while different
initializations converge to similar ultimate values, we can never
realistically obtain the global optimum. There are a broad range of
techniques that may offer better approximations to the global optimum
\autocite{Wanga2012,FritzJonathan} and future work could adapt them to
the problem we addressed here.

@@

Our model is a extension of infinite stochastic block models
\autocite{Kemp2006a,Xu2006} to incorporate spatial constraints.

\todo{infinite subset relaitonal model has good summaries}

@@

\subsection{Extension to multiple graphs}
The model can handle multiple graphs $R^q$ simultaneously with a shared clustering by simply considering the product of 
\begin{equation}
  p(\vec{c}, \{\eta^q\}, \{\theta^q\} | \{R^q\} ) \propto \prod_q \Bigg(\prod_{i, j} p(R^q(e_i, e_j) | f(d(e_i, e_j) | \eta^q_{c_ic_j}), \theta^q) \prod_{m, n} p(\eta^q_{mn} | \theta^q)  p(\theta^q) \Bigg) p(\vec{c} | \alpha) p(\alpha) 
\end{equation}

\subsection{Microprocessor Netlist annotation}
\begin{figure}
  \centering 
    \centerline{\includegraphics[width=183mm]{mos6502/sixrelation.xysregs.ldfl.circos.4.R2.ai}}
\end{figure}

\subsection{Hyperprior grids}

For the mouse retina Logistic-Distance Bernoulli model, we gridded $\mu^{hp}$ and $\lambda^{hp}$ into 40 $\log_{10}$-spaced points 1.0 and 80. $p_{min} \in [0.001, 0.005, 0.01]$ and $p_{max} \in [0.80, 0.90, 0.95, 0.99$.

For the c. elegans data with the Logistic Distance poisson model, we gridded $\mu_{hp$ and $\lambda$ into 20 $\log_{10}$-spaced points bween 0.2 and 2.0, and the $ratescale^{hp}$ parameter into 20 $\log_{10}$-spaced points between 2.0 and 20.0. We globally set $rate_{min}=0.01$.
